---
output:
  md_document:
    variant: markdown_github
---

Starting with the basics in this notebook based on UvA's [Tutorial 2](https://uvadlc-notebooks.readthedocs.io/en/latest/tutorial_notebooks/tutorial2/Introduction_to_PyTorch.html)

```{r setup}
library(reticulate)
library(here)
```

Here's my python configuration for this project.

```{r}
py_config()
```

I wonder which of the standard Python libraries I have already? Let's check.

```{python}
import os
import math
import numpy as np
import time

## Imports for plotting
import matplotlib.pyplot as plt
from matplotlib.colors import to_rgba
import seaborn as sns
sns.set()

## Progress bar
from tqdm.notebook import tqdm
```

All imported with no issues. Very nice. Thanks, conda. \<3

# The Basics of PyTorch

Now we'll import torch.

```{python}
import torch
print("Using torch", torch.__version__, "with CUDA", torch.cuda_version)
```

## Tensors

OK, now we'll move on to some `tensor` stuff. In ML/DL context, tensors are basically multidimensional arrays. If you want to go down a rabbit hole about why they're called tensors when tensors are something very specific in mathematics, then you can [start here](https://stats.stackexchange.com/a/198395). ðŸ‡

```{python}
x = torch.Tensor(2, 3, 4)
print(x)
```

Oooh, neat. Apparently memory is allocated when we use `torch.Tensor` but the values it initializes with are those that have already been in memory. I like that (I don't know why, but I do).

Other ways to specify values for tensors:

-   `torch.zeros` - values filled with zeros\
-   `torch.ones` - values filled with ones\
-   `torch.rand` - values filled with samples drawn from a uniform distribution between 0 and 1\
-   `torch.randn` - values filled with samples drawn from a normal distribution with mean 0 and variance 1\
-   `torch.arange` - values are filled with N, N+1, N+2, ..., M. The step can be non-integer, e.g.

```{python}
torch.arange(0, 1, 0.05)
```

We can use nested lists to specify the elements of a tensor directly.

```{python}
x = torch.Tensor([[1, 2], [3, 4]])
print(x)
```

The methods to obtain the shape of a tensor are `size` and `shape`.

```{python}
print("Shape:", x.shape)
print("Size:", x.size())
```

## Tensor to Numpy and vice versa

We use the `from_numpy` to go from a numpy array to a tensor.

```{python}
np_arr = np.array([[1, 2], [3, 4]])
tensor = torch.from_numpy(np_arr)

print(tensor)
```

And `.numpy()` to go from a PyTorch tensor to a numpy array.

```{python}
tensor = torch.arange(4)
np_arr = tensor.numpy()

print("I'm a", tensor)
print("And here I am as an Numpy array: ", np_arr)
```

An important note from the tutorial:

> The conversion of tensors to numpy require the tensor to be on the CPU, and not the GPU (more on GPU support in a later section). In case you have a tensor on GPU, you need to call .cpu() on the tensor beforehand. Hence, you get a line like np_arr = tensor.cpu().numpy().

## Operations

Check the [PyTorch docs](https://pytorch.org/docs/stable/tensors.html#) for the full set of tensor operations available.

We can either create new tensors using operations or use methods to perform in-place operations that will modify the tensor. These usually have an underscore postfix.

```{python}
x1 = torch.rand(2, 3)
x2 = torch.rand(2, 3)
print("x1 before:\n", x1)

# Now add to x1
x1.add_(x2)

print("x1 after:\n", x1)
```

To re-shape tensors we can use `view` and `permute` operations.

View will add on a row and column shape.

```{python}
x = torch.arange(6)
print(x, "\nshapes into")
x.view(2, 3)
```

Permute swaps the dimensions specified.

```{python}
x = x.view(2, 3)
x.permute(0, 1)
```

You have to make sure you permute using the same number of dimensions as the tensor.

Other operations covered in the tutorial are for matrix multiplication.

-   `torch.matmul` - performs matrix product based on the dimensions of the tensors. If both are 2-dim then it will be a standard matrix product. Higher dimensional inputs will use [broadcasting](https://pytorch.org/docs/stable/notes/broadcasting.html#broadcasting-semantics). Can also be written as `a @ b`, which is similar to numpy.

-   `torch.mm` - matrix product over two matrices but doesn't support broadcasting.

-   `torch.bmm` - Performs the matrix product with a support batch dimension. If the first tensor is of shape (b, n, m), and the second tensor (b, m, p), the output is of shape (b, n, p). Basically it uses the first dimension as an index and multiplies the matrices of dimension (n,m) from the first argument by the matrices of dimension (n,p) from the second argument at each index.

-   `torch.einsum` - Performs matrix multiplications and more (i.e. sums of products) using the Einstein summation convention.

Shneat. I gotta try some of these batch mm examples to make sense of that.
